Reasoning About a Capability Machine with Local Capabilities

### 2017-09-28 esop-2018/paper.tex
Improve alignment in a figure.
Fix affiliation.
Save some space by using wrapfig and playing with figure placement.
Save some more space.

\usepackage{cite}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{thmtools}
\usepackage{stmaryrd}
% \usepackage[numbers]{natbib}
\usepackage{url}
\usepackage{array}
\usepackage{arydshln}
\usepackage{ifthen}
\usepackage{ifpdf}
\usepackage{verbatim}
\usepackage{mathpartir}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{multicol}
\lstset{
  basicstyle=\footnotesize\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape
}

@@

% Authors and affiliation
\author{Lau~Skorstengaard\inst{1} \and Dominique~Devriese\inst{2} \and Lars~Birkedal\inst{1}}
\institute{Aarhus University\email{\{lau,birkedal\}@cs.au.dk}
\and iMinds-DistriNet, KU Leuven\email{dominique.devriese@cs.kuleuven.be}}

@@

\begin{figure}
\[\arraycolsep=1.4pt
  \begin{array}{rlrcl l rlrcl}
    \addr   &\in& \Addrs &\isdef& \nats &\phantom{mak}                                & r    &\in & \RegName     &\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
    w &\in&\Words &\isdef& \ints + \Caps &                                                   & \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words \\
    \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid  & & m    &\in & \Heaps       &\isdef& \Addrs \rightarrow \Words \\
    & & & &\exec\mid \entry\mid \rwx\mid \rwlx &                                             & \Phi &\in & \ExecConfs   &\isdef& \Regs \times \Heaps \\
    \gl&\in&\Globals & ::= &\glob \mid \local &                                              & \ms  &\in & \MemSegments &::= & \Addrs \parfun \Words \\
    &&\Confs &::=& \multicolumn{6}{l}{\ExecConfs + \{\failed \} + \{\halted\} \times \Heaps}\\
     && \Caps  &::= & \multicolumn{6}{l}{\left\{
                         ((\perm,\gl),\start,\addrend,\addr)\mid
                         \start,\addr\in\Addrs,\addrend \in
                         \Addrs\cup \{\infty\}\right\}}
  \end{array}
\]
  \begin{equation*}
  \begin{array}{rcl}
    \rv    &\in& \ints + \RegName \\
    i      &::=& 
                 \jmp{\lv} \mid 
                 \jnz{\lv}{\rv} \mid
                 \move{\lv}{\rv} \mid 
                 \load{\lv}{\hv} \mid 
                 \store{\hv}{\rv} \mid \plus{\lv}{\rv}{\rv} \mid 
                 \minus{\lv}{\rv}{\rv} \mid \\
          &   &  \lt{\lv}{\rv}{\rv} \mid
                \lea{\lv}{\rv} \mid 
                \restricttwo{\lv}{\rv} \mid
                 \subseg{\lv}{\rv}{\rv} \mid  
                 \isptr{\lv}{\rv} \mid 
                 \getl{\lv}{\lv} \mid \\
          &   &  \getp{\lv}{\lv} \mid 
                 \getb{\lv}{\lv} \mid
                 \gete{\lv}{\lv} \mid 
                 \geta{\lv}{\lv} \mid 
                 \fail \mid
                 \halt 
  \end{array}
\end{equation*}
  \caption{The syntax of our capability machine assembly language.}
  \label{fig:syntax}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}[main node/.style={}]
    \node[main node] (7) {$\rwlx$};
    \node[main node] (8) [below left of=7] {$\readwritel$};
    \node[main node] (1) [below right of=7] {$\rwx$};
    \node[main node] (2) [below right of=1] {$\exec$};
    \node[main node] (3) [below right of=2] {$\entry$};
    \node[main node] (4) [below left of=1] {$\readwrite$};
    \node[main node] (5) [below right of=4] {$\readonly$};
    \node[main node] (6) [below right of=5] {$\noperm$};

@@

  \caption{Permission hierarchy}
  \label{fig:perm-hier}
\end{figure}

@@

\subsection{Worlds}
\begin{figure}
  \centering
  \includegraphics{w11}
  \caption{An illustration of local/global capabilities and temporary/permanent
    regions. The table represents a memory and the colored fields are regions
    governing a memory range. Global capabilities cannot depend on temporary
    regions.}
  \label{fig:cap-world}
\end{figure}
% \dominique{Perhaps we could first try to convey the intuition of a permanent
%   world with a temporary overlay, with global capabilities ignoring the overlay.
%   Perhaps we could add a depiction of the world to Figure 5?}
%\todo[inline]{Find a good place to describe monotonicity of
%interpretation function.}
% \lau{this may not be an important enough point to include it.}

@@

\begin{figure}[htbp]
  \centering
  \begin{align*}
  & \readCond{}(\gl)(W) =  \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{array}{l}
       \exists r \in \var{localityReg}(g,W) \ldotp \\
\quad   \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{array}
    \right\}\\
  & \writeCond{}(\iota,\gl)(W) =  \left\{
    \npair{(\start,\addrend)}
    \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & W(r) \text{ is address-stratified and} \\[-0.2cm]
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
    \end{aligned} \right\}\\
  & \execCond{}(\gl)(W) = 
    \left\{
      \npairP{(\var{P}, \start,\addrend)}
     \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W, a \in [\start,\addrend], \perm \in \var{P} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
  & \entryCond{}(\gl)(W) = 
    \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{array}{l}
      \forall n' < n \ldotp \forall W' \future W \ldotp\\
      \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{array}
    \right\} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad\;\; \text{and } \gl = \glob \Rightarrow \future = \futurestr
  \end{align*}
\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
}
%subsection introduction
% \dominique{I would try to motivate the LR defs by saying that the LR is
%   constructed to express the FTLR and the FTLR says that if a piece of code only
%   has access to values respecting the invariants in a world, then it will
%   respect them itself too.} 
The logical relation defines semantically when values, program counters, and
configurations are safe according to the intuitions described in
Section~\ref{sec:formalizing-guarantees}. The definition is found in
\figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds} and we provide some
explanations in the following paragraphs. We omit some definitions and explain
them only verbally, but precise definitions can be found in the technical
appendix~\cite{technical_appendix}.

@@

\begin{figure}[t]
  \centering
  \subfloat{\usebox{\testit}}
  \subfloat{\usebox{\progfthree}}
  \caption{test}
\end{figure}
\end{comment}
\begin{figure}[t]
  \centering
  \begin{minipage}[t]{5.05cm}
  \begin{lstlisting}
g1:  malloc $r_2$ 1
     store $r_2$ 0
     move $\pcreg$ $r_3$
     lea $r_3$ $\var{offset}$
     crtcls $[(x, r_2)]$ $r_3$
     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
     jmp $r_0$
f4:  reqglob $r_1$
     prepstk $r_\stk$
     store $x$ 0
     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
     store $x$ 1
     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
     load $r_1$ $x$
     assert $r_1$ 1
     mclear $r_\stk$
     rclear $\RegName \setminus \{r_0,\pcreg\}$
     jmp $r_0$
\end{lstlisting}
  \end{minipage}
  \begin{minipage}[t]{3.4cm}
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \end{minipage}
  \caption{ Two programs that rely on well-bracketedness of
    \texttt{scall}s to function correctly. $\var{offset}$ is the
    offset to \texttt{f4}.}
  \label{fig:prog-f3-and-g1}
\end{figure}

>>>>

\usepackage{cite}
\usepackage{wrapfig}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{thmtools}
\usepackage{stmaryrd}
% \usepackage[numbers]{natbib}
\usepackage{url}
\usepackage{array}
\usepackage{arydshln}
\usepackage{ifthen}
\usepackage{ifpdf}
\usepackage{verbatim}
\usepackage{mathpartir}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{multicol}
\lstset{
  basicstyle=\footnotesize\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape
}

@@

% Authors and affiliation
\author{Lau~Skorstengaard\inst{1} \and Dominique~Devriese\inst{2} \and Lars~Birkedal\inst{1}}
\institute{Aarhus University\email{\{lau,birkedal\}@cs.au.dk}
\and imec-DistriNet, KU Leuven\email{dominique.devriese@cs.kuleuven.be}}

@@

\begin{figure}[p]
\[\arraycolsep=1.4pt
  \begin{array}{rlrcl l rlrcl}
    \addr   &\in& \Addrs &\isdef& \nats &\phantom{mak}                                & r    &\in & \RegName     &\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
    w &\in&\Words &\isdef& \ints + \Caps &                                                   & \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words \\
    \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid  & & m    &\in & \Heaps       &\isdef& \Addrs \rightarrow \Words \\
    & & & &\exec\mid \entry\mid \rwx\mid \rwlx &                                             & \Phi &\in & \ExecConfs   &\isdef& \Regs \times \Heaps \\
    \gl&\in&\Globals & ::= &\glob \mid \local &                                              & \ms  &\in & \MemSegments &::= & \Addrs \parfun \Words \\
    &&\Confs &::=& \multicolumn{6}{l}{\ExecConfs + \{\failed \} + \{\halted\} \times \Heaps}\\
     && \Caps  &::= & \multicolumn{6}{l}{\left\{
                         ((\perm,\gl),\start,\addrend,\addr)\mid
                         \start,\addr\in\Addrs,\addrend \in
                         \Addrs\cup \{\infty\}\right\}}
  \end{array}
\]
  \begin{equation*}
  \begin{array}{rcl}
    \rv    &\in& \ints + \RegName \\
    i      &::=& 
                 \jmp{\lv} \mid 
                 \jnz{\lv}{\rv} \mid
                 \move{\lv}{\rv} \mid 
                 \load{\lv}{\hv} \mid 
                 \store{\hv}{\rv} \mid \plus{\lv}{\rv}{\rv} \mid 
                 \minus{\lv}{\rv}{\rv} \mid \\
          &   &  \lt{\lv}{\rv}{\rv} \mid
                \lea{\lv}{\rv} \mid 
                \restricttwo{\lv}{\rv} \mid
                 \subseg{\lv}{\rv}{\rv} \mid  
                 \isptr{\lv}{\rv} \mid 
                 \getl{\lv}{\lv} \mid \\
          &   &  \getp{\lv}{\lv} \mid 
                 \getb{\lv}{\lv} \mid
                 \gete{\lv}{\lv} \mid 
                 \geta{\lv}{\lv} \mid 
                 \fail \mid
                 \halt 
  \end{array}
\end{equation*}
  \caption{The syntax of our capability machine assembly language.}
  \label{fig:syntax}
\end{figure}

\begin{wrapfigure}{R}{.3\textwidth}
  \centering
  \begin{tikzpicture}[main node/.style={}]
    \node[main node] (7) {$\rwlx$};
    \node[main node] (8) [below left of=7] {$\readwritel$};
    \node[main node] (1) [below right of=7] {$\rwx$};
    \node[main node] (2) [below right of=1] {$\exec$};
    \node[main node] (3) [below right of=2] {$\entry$};
    \node[main node] (4) [below left of=1] {$\readwrite$};
    \node[main node] (5) [below right of=4] {$\readonly$};
    \node[main node] (6) [below right of=5] {$\noperm$};

@@

  \caption{Permission hierarchy}
  \label{fig:perm-hier}
\end{wrapfigure}

@@

\subsection{Worlds}
\begin{figure}[tbp]
  \centering
  \resizebox{6cm}{!}{\includegraphics{w11}}

  \caption{An illustration of local/global capabilities and temporary/permanent
    regions. The table represents a memory and the colored fields are regions
    governing a memory range. Global capabilities cannot depend on temporary
    regions.}
  \label{fig:cap-world}
\end{figure}
% \dominique{Perhaps we could first try to convey the intuition of a permanent
%   world with a temporary overlay, with global capabilities ignoring the overlay.
%   Perhaps we could add a depiction of the world to Figure 5?}
%\todo[inline]{Find a good place to describe monotonicity of
%interpretation function.}
% \lau{this may not be an important enough point to include it.}

@@

\begin{figure}[htbp]
  \centering
  \begin{align*}
   \readCond{}(\gl)(W) &=  \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{array}{l}
       \exists r \in \var{localityReg}(g,W) \ldotp \\
\quad   \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{array}
    \right\}\\
   \writeCond{}(\iota,\gl)(W) &=  \left\{
    \npair{(\start,\addrend)}
    \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & W(r) \text{ is address-stratified and} \\[-0.2cm]
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
    \end{aligned} \right\}\\
   \execCond{}(\gl)(W) &= 
    \left\{
      \npairP{(\var{P}, \start,\addrend)}
     \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W, a \in [\start,\addrend], \perm \in \var{P} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
   \entryCond{}(\gl)(W) &= 
    \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{array}{l}
      \forall n' < n \ldotp \forall W' \future W \ldotp\\
      \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{array}
    \right\} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad\;\; \text{and } \gl = \glob \Rightarrow \future = \futurestr
  \end{align*}
\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
}
%subsection introduction
% \dominique{I would try to motivate the LR defs by saying that the LR is
%   constructed to express the FTLR and the FTLR says that if a piece of code only
%   has access to values respecting the invariants in a world, then it will
%   respect them itself too.} 
The logical relation defines semantically when values, program counters, and
configurations are safe according to the intuitions described in
Section~\ref{sec:formalizing-guarantees}. The definition is found in
\figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds} and we provide some
explanations in the following paragraphs. We omit some definitions and explain
them only verbally, but precise definitions can be found in the technical
appendix~\cite{technical_appendix}.

@@

\begin{figure}[t]
  \centering
  \subfloat{\usebox{\testit}}
  \subfloat{\usebox{\progfthree}}
  \caption{test}
\end{figure}
\end{comment}
\begin{figure}[t]
  \centering
  \begin{minipage}[t]{.37\linewidth}
  \begin{lstlisting}
g1:  malloc $r_2$ 1
     store $r_2$ 0
     move $\pcreg$ $r_3$
     lea $r_3$ $\var{offset}$
     crtcls $[(x, r_2)]$ $r_3$
     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
     jmp $r_0$
f4:  reqglob $r_1$
     prepstk $r_\stk$
     ${\tiny\itshape\text{(continues in next column)}}$
\end{lstlisting}
  \end{minipage}%
  \begin{minipage}[t]{.37\linewidth}
\begin{lstlisting}
     ${\tiny\itshape\text{(continued from previous column)}}$
     store $x$ 0
     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
     store $x$ 1
     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
     load $r_1$ $x$
     assert $r_1$ 1
     mclear $r_\stk$
     rclear $\RegName \setminus \{r_0,\pcreg\}$
     jmp $r_0$
\end{lstlisting}
  \end{minipage}%
  \begin{minipage}[t]{.25\linewidth}
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \end{minipage}
  \caption{ Two programs that rely on well-bracketedness of
    \texttt{scall}s to function correctly. $\var{offset}$ is the
    offset to \texttt{f4}.}
  \label{fig:prog-f3-and-g1}
\end{figure}
